\ieeetitle

### 2016-12-30 report/framework.tex
Add figures as standalone pdfs and fix issues
*Issues were mentionned by Hugues

\section{EpToTester}
\label{sec:epto}
\begin{figure}[htp]
	\includegraphics[width=\linewidth]{figures/architecture.pdf}
	\caption{\epto architecture \autocite{matos2015epto}.}
	\label{fig:epto-architecture}
\end{figure}
EpTOTester is a practical implementation of \epto designed to assess the claims made in \autocite{matos2015epto}. Although the implementation was written with benchmarking in mind, the code could be adapted to be used in a real application with only minimal changes to the sources.
\subsection{Architecture}
\autoref{fig:epto-architecture} illustrates the architecture of a single replica. An application extending our Application class can \epto-broadcast and \epto-deliver events. Events broadcasted to the Dissemination component are sent over the network every $\delta$ period. Every ball received from the network is unwrapped and its events are analyzed by the Dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL). They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order. The order is based on the timestamp of the events given by the logical clock and their broadcaster ID in case of a tie. The network layer communication is done using UDP and a PSS CYCLON to obtain a random view of peers. To obtain an initial random view, we contact an independent tracker that keeps track of dead and alive nodes in the system. We want to emphasize that the tracker is not required. In practice it works well, but using a DHT is certainly a possibility.
\subsection{Implementation}
We implement the data sent as randomly generated UUIDs. We implement our own PSS CYCLON operating on its own port. We use \jgroups 3.6.11. When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers. This is not a problem as in a real WAN \jgroups could not rely on MULTICAST. Finally, EpToTester is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
\subsection{Deployment}
\begin{figure}[htp]
	\centering
	\input{figures/architecture.tex}
	\vspace{-2mm} 
	\caption{EpTOTester architecture}
	\vspace{-2mm} 
	\label{fig:complete-architecture}
\end{figure}
To deploy \epto effectively we need two different Docker Swarm services. One containing the tracker and one containing all \epto replicas. Both of these services use the same network overlay to communicate. This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12. This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications. Every benchmark is started through a Python script available on the master mode. This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results. All \epto parameters are customizable by using options provided through this script.
Gradle is used to automate the image building in the python scripts. Finally, a script is provided to push the images to a repository accessible to the remote cluster and to push the benchmarking script on the master node. \jgroups deployment is much the same. The entire EpTOTester architecture is shown in \autoref{fig:complete-architecture}. \jt{The figure is inspired from the one in ErasureBench. How should I indicate this?}
\subsection{Fault Injection}
Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}. The synthetic churn provided by it was improved to support adding and removing nodes at the same time.

>>>>

\section{EpToTester}
\label{sec:epto}
\begin{figure}[htp]
	\includegraphics[width=\linewidth]{figures/architecture.pdf}
	\caption{\epto architecture \autocite{matos2015epto}.}
	\label{fig:epto-architecture}
\end{figure}
\eptotester is a practical implementation of \epto designed to assess the claims made in \autocite{matos2015epto}. Although the implementation in written with benchmarking in mind, the code could be adapted for real applications with only minimal changes to the sources. \eptotester is open-sourced on Github\footnote{\href{https://github.com/jocelynthode/EpTOTester}{https://github.com/jocelynthode/EpTOTester}}.
\subsection{Architecture}
\autoref{fig:epto-architecture} illustrates the architecture of a single replica. An application extending our Application class can \epto-broadcast and \epto-deliver events. Events broadcasted to the Dissemination component are sent over the network every $\delta$ period. Every ball received from the network is unwrapped and its events are analyzed by the Dissemination component to find out whether they need to be propagated further or not according to their Time To Live (TTL). They are then sent to the ordering component so that \epto can determine whether to deliver these events or not and in which order. The order is based on the timestamp of the events given by the logical clock and their broadcaster ID in case of a tie. The network layer communication is done using UDP and a PSS CYCLON to obtain a random view of peers. To obtain an initial random view, we contact an independent tracker that keeps track of dead and alive nodes in the system. We want to emphasize that the tracker is not required. In practice it works well, but using a DHT is certainly a possibility.
\subsection{Implementation}
We implement the payload as randomly generated UUIDs. We implement our own PSS CYCLON operating on its own port.  We compare \epto to the deterministic total order algorithm provided by \jgroups. We use \jgroups 3.6.11. When implementing \jgroups we use TCPGOSSIP provided by the \jgroups library instead of the traditional MULTICAST option to coordinate peers. This is not a problem as in a real WAN \jgroups could not rely on MULTICAST.
\subsection{Deployment}
 \begin{figure}[htp]
 	\centering
 	\includegraphics[width=\linewidth]{figures/complete-architecture.pdf}
 	\vspace{-2mm} 
 	\caption[Caption]{\eptotester architecture\footnotemark}
 	\vspace{-2mm} 
 	\label{fig:complete-architecture}
 \end{figure}
\footnotetext{This figure is partially inspired from a figure in \autocite{vaucher2016erasure}}
To deploy \epto effectively we need two different Docker Swarm services: one containing the tracker and one containing all \epto replicas. Both of these services use the same network overlay to communicate. This is achieved using Docker and especially the new Docker Swarm introduced in Docker 1.12. This lets us have a unified way of deploying our benchmarks locally or remotely on vastly different clusters with minimal modifications. Every benchmark is started through a Python script available on the master node. This script handles everything from starting benchmarks, scaling the cluster during a churn and collecting results. All \epto parameters are customizable by using options provided through this script.
Gradle\footnote{\href{https://gradle.org/}{https://gradle.org/}} is used to automate the project building. Finally, a script is provided to push the images to a repository accessible to the remote cluster and to push the benchmarking script on the master node. \jgroups deployment is much the same. The entire \eptotester architecture is shown in \autoref{fig:complete-architecture}.
\subsection{Fault Injection}
Our framework supports the ability to inject synthetic and real traces thanks to the work done in \autocite{vaucher2016erasure}. The synthetic churn provided by it was improved to support adding and removing nodes at the same time.

### 2016-12-30 report/introduction.tex
Add figures as standalone pdfs and fix issues
*Issues were mentionned by Hugues

Creating an algorithm providing scalability, integrity and validity, along with a total ordering of the events through all peers in a distributed system has been one the hot topics in distributed systems research for many years. One of the recently designed algorithms on this topic is \epto \autocite{matos2015epto}. \epto is an algorithm that claims to provide integrity, validity and total order in a large-scale distributed system. In addition, \epto is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for dynamic large-scale distributed systems.

@@

There are many other algorithms for disseminating and ordering events in a distributed system. There are some deterministic algorithms, which guarantee total order, agreement or other strong properties. Unfortunately, these types of algorithms are not scalable enough to be used in large-scale distributed systems \autocites[]{defago2004total}[]{lamport1978time}.

@@

The problem with existing deterministic total ordering protocols is that they need some sort of agreement between all peers in the system. This causes a massive amount of network traffic and overhead on the system.

@@

explicitly maintain a group and have access to a failure detector \autocites[]{chandra1996weakest}[]{chandra1996unreliable}. Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the system and thus limits the algorithm's scalability. \hm{The first three paragraphs use the word system a lot. You could rework this a little bit.}

@@

As an alternative to deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach \hm{Write [6-13] instead.} \autocites []{birman1999bimodal}[]{carvalho2007emergent}[]{demers1987epidemic}[]{eugster2003lightweight}[]{felber2002probabilistic}[]{hayden1996probabilistic}[]{kim2004gossip}[]{Koldehofe02simplegossiping}. These algorithms guarantee the dissemination of events in the system with high probability. This way, there is no need for failure detectors and redundant traffic, making these algorithms highly scalable. As these algorithms focus on reliability of dissemination, they often have to ignore other properties such as total ordering.

@@

This is where \epto shines. \hm{What does ``where'' refer to? Not clear.} \epto, by mixing these two types of algorithms, \hm{Not sure I agree here. What two types? Not clear. Epto is not probabilistic and deterministic. Please rephrase.} provides total order along with scalability, validity and integrity. 
\epto consists of two distinct parts. The first part is probabilistic dissemination. \epto guarantees that all peers will receive an event with arbitrarily high probability. The second part of \epto is deterministic ordering. Once peers have received all events, they deterministically order them using the events timestamp, and in case of a tie use the broadcaster ID of the events. \hm{Write component instead of part. Fix before and after.}

@@

To model the first part, \epto is using a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}. The \textit{balls-and-bins} problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin uniformly at random and independently from other balls. In this scenario, one of the natural questions that comes to mind is: what is the minimum number of balls that should be thrown, so that every bin has at least one ball with high probability?

@@

The next sections present our benchmarking solution and the results obtained. \autoref{sec:ordering} presents the different kinds of ordering. \autoref{sec:definitions} defines the terms used in the paper. \autoref{sec:epto} presents \epto and the architecture of our implementation. \autoref{sec:evaluation} shows and explains the results obtained as well as the limitations our project faced. We then present possible future work in \autoref{sec:future-work}. We finally conclude in \autoref{sec:conclusion}. \hm{Write Section X instead of section X. Scan all document.}

>>>>

Creating an algorithm providing scalability, integrity and validity, along with a total ordering of the events through all peers in a distributed network has been one the hot topics in distributed computing research for many years. One of the recently designed algorithms on this topic is \epto \autocite{matos2015epto}. \epto is an algorithm that claims to provide integrity, validity and total order in a large-scale distributed system. Furthermore, \epto is designed to work without a global clock, removing the need to synchronize clocks precisely on every peer and thus is well-suited for these kind of systems.

@@

There are many other algorithms for disseminating and ordering events in a distributed network. There are deterministic algorithms, which guarantee total order, agreement or other strong properties. Unfortunately, these types of algorithms are not scalable enough to be used in large-scale distributed systems \autocites[]{defago2004total}[]{lamport1978time}.

@@

The problem with existing deterministic total ordering protocols is that they need some sort of agreement between all peers in the network. This causes a massive amount of network traffic and overhead on the system.

@@

explicitly maintain a group and have access to a failure detector \autocites[]{chandra1996weakest}[]{chandra1996unreliable}. Due to faults and churn in large-scale distributed systems, the failure detector turns into a bottleneck for the structure and thus limits the algorithm's scalability. \hm{The first three paragraphs use the word system a lot. You could rework this a little bit.} \jt{I tried replacing some 'system' by synonyms such as 'network' or 'structure'.}

@@

As an alternative to deterministic algorithms, there are probabilistic algorithms, focusing on scalability and resiliency against failures using a probabilistic dissemination approach  \autocite{birman1999bimodal,carvalho2007emergent,demers1987epidemic,eugster2003lightweight,felber2002probabilistic,hayden1996probabilistic,kim2004gossip,Koldehofe02simplegossiping}. These algorithms guarantee the dissemination of events in the system with high probability. This way, there is no need for failure detectors and redundant traffic, making these algorithms highly scalable. As these algorithms focus on reliability of dissemination, they often have to ignore other properties such as total ordering.

@@

\epto claims to solve these seemingly irreconcilable differences. By using a probabilistic dissemination together with deterministic ordering, \epto provides total order along with scalability, validity and integrity. Thus, \epto consists of two distinct components. \epto's dissemination component guarantees that all peers will receive an event with arbitrarily high probability. Once peers have received all events, \epto's ordering component orders them using the events timestamp, and in case of a tie use the broadcaster ID of the events.

@@

To model the first component, \epto is using a \textit{balls-and-bins} approach \autocite{Koldehofe02simplegossiping}. The \textit{balls-and-bins} problem is a basic probabilistic problem: consider \textit{n} balls and \textit{m} bins where we consequently throw balls into a bin uniformly at random and independently from other balls. In this scenario, one of the natural questions that comes to mind is: what is the minimum number of balls that should be thrown, so that every bin has at least one ball with high probability?

@@

The next sections present our benchmarking solution and the results obtained. Section \ref{sec:ordering} presents the different kinds of ordering. Section \ref{sec:definitions} defines the terms used in the paper. Section \ref{sec:epto} presents \epto and the architecture of our implementation. Section \ref{sec:evaluation} shows and explains the results obtained as well as the limitations our project faced. We then present possible future work in Section \ref{sec:future-work}. We finally conclude in Section \ref{sec:conclusion}.

### 2016-12-30 report/ordering.tex
Add figures as standalone pdfs and fix issues
*Issues were mentionned by Hugues

\section{Ordering Algorithms}
\label{sec:ordering}
\jt{Do we still want this section?}\hm{Absolutely!}
Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system. When there are separated processes that can only communicate through messages, you cannot easily order these messages.
Therefore we need ordering algorithms to overcome this problem.
\par
We have two types of ordering algorithms \autocite{lamport1978time}: the partial order algorithms and the total order algorithms.
\subsection{Partial Order Algorithms}
Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
\begin{itemize}
	\item Reflexivity: $a \leq a$ for all $a \in S$.
	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
\end{itemize}

>>>>

\section{Ordering Algorithms}
\label{sec:ordering}
Distributed systems and centralized systems alike, need to preserve the temporal order of events produced by concurrent processes in the system. When there are separated processes that can only communicate through messages, you cannot easily order these messages.
Therefore we need ordering algorithms to overcome this problem.
\par
We have two types of ordering algorithms \autocite{lamport1978time}: the partial order algorithms and the total order algorithms.
\subsection{Partial Order Algorithms}
Assuming S is partially ordered under $\leq$, then the following statements hold for all a, b and c in S:
\begin{itemize}
	\item Reflexivity: $a \leq a$ for all $a \in S$.
	\item Antisymmetry: $a \leq b$ and $b \leq a$ implies $a=b$ .
	\item Transitivity: $a \leq b$  and $b \leq c$  implies $a \leq c$.
\end{itemize}
